---
title: "Exporting Review Tables to LaTeX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting Review Tables to LaTeX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`reviewTable()` can return a LaTeX representation of the summary table for
inclusion in a manuscript. Two arguments control the LaTeX workflow:

- `latex = TRUE` returns the table as LaTeX code via `gt::as_latex()` instead
  of a `gt` object.
- `cite = TRUE` treats values in the study ID column as BibTeX reference keys
  and wraps each row's study list in `\cite{...}`, so citations render
  correctly when the document is compiled.

```{r setup}
library(litReview)
data(studies)
```

## Plain LaTeX output

Pass `latex = TRUE` to get LaTeX code you can paste into a `.tex` document or
print inside an R Markdown chunk with `results = "asis"`:

```{r latex-plain}
reviewTable(studies, Design, latex = TRUE)
```


## Study IDs as BibTeX keys

If `StudyID` matches the keys in your `.bib` file (e.g. `smith2020`,
`doe2019`), set `cite = TRUE` so the `Studies` column is wrapped in
`\cite{...}`:

```{r latex-cite}
reviewTable(studies, Design, latex = TRUE, cite = TRUE)
```

The `cite` flag also works without `latex = TRUE` if you want a `gt` preview
of the cited form:

```{r gt-cite}
reviewTable(studies, Design, cite = TRUE, study_id = "bibKey")
```

## Embedding in a LaTeX or R Markdown document

In an `.Rnw` (Sweave) or LaTeX-output R Markdown file, print the result with
`results = "asis"` so the LaTeX code is emitted verbatim:

````
```{r, results = "asis"}
reviewTable(studies, Design, latex = TRUE, cite = TRUE)
```
````

Make sure your preamble loads a citation package (e.g. `\usepackage{natbib}`
or `\usepackage{biblatex}`) and that the keys used in `StudyID` exist in the
bibliography.

## Custom ID column

If your study IDs live in a column other than `StudyID`, pass it via
`study_id`:

```{r custom-id}
df <- data.frame(
  RefKey = c("smith2020", "doe2019", "lee2021", "smith2020"),
  Design = c("RCT", "Cohort", "RCT", "Case-control"),
  stringsAsFactors = FALSE
)
reviewTable(df, Design, study_id = RefKey, latex = TRUE, cite = TRUE)
```
