---
title: "Alluvial diagrams"
description: >
  Every configuration option of reviewAlluvial(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
has_ggalluvial <- requireNamespace("ggalluvial", quietly = TRUE)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewAlluvial()` shows how studies flow across two or more categorical
columns: each study traces a path through the strata, and the ribbons reveal
which category combinations co-occur. It is the tool for spotting patterns like
"which designs measure which outcomes." This article walks through **every
argument** of

```r
reviewAlluvial(data, cols, sep = "\r\n", study_id = StudyID, colors = PALETTE,
               base_size = 12, na.rm = TRUE, na_label = "Not reported",
               labels = c("none", "prop", "count", "both"), flow_labels = FALSE,
               flow_alpha = 0.25, stratum_width = 0.5, axis_labels = NULL)
```

It requires the **ggalluvial** package.

## Default

Pass the data frame and a character vector of column names. Each axis is one
column; the boxes (strata) are its categories, and the ribbons connecting them
are the studies.

```{r default, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"))
```

## `cols`: the columns to trace

`cols` is a **character vector of two or more column names**, one axis per
name, drawn left to right. A two-column diagram links designs to outcomes:

```{r cols-two, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"))
```

Add a third column and every study now traces a path across three axes.
Multi-value columns (`Outcome`, `AgeGroup`) are split first, so a study with
several values contributes a flow to each:

```{r cols-three, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome", "AgeGroup"))
```

## `sep`: multi-value separator

Some cells hold several values. In `studies`, `Outcome` and `AgeGroup` use
newline separators (`"\r\n"`, the default), so each value spawns its own flow.
If your data uses a different delimiter, set `sep`. Here we rebuild a
semicolon-separated column to demonstrate:

```{r sep-custom, fig.height = 5, eval = has_ggalluvial}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewAlluvial(studies_semi, c("Design", "Outcome"), sep = "; ")
```

## `study_id`: the identifier column

`study_id` names the column that identifies one study (default `StudyID`). It
defines what a single flow is: rows sharing an ID are one study, so multi-value
cells fan out from the same origin. Any unique identifier works — here we use
`Author`:

```{r study-id, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), study_id = Author)
```

## `colors`: the stratum palette

Strata are filled from a palette vector (default `PALETTE`). Print it to see the
default colors:

```{r colors-palette-show}
PALETTE
```

Pass your own vector to recolor the strata (recycled if shorter than the number
of categories):

```{r colors-custom, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"),
               colors = c("#4E79A7", "#F28E2B", "#E15759", "#76B7B2",
                          "#59A14F", "#EDC948"))
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), base_size = 18)
```

## Missing data: `na.rm` and `na_label`

By default `na.rm = TRUE` drops studies with a missing (or empty) value in any
of the plotted columns. We use `RiskOfBias` and `FundingSource` — the latter
has missing values:

```{r na-default, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("RiskOfBias", "FundingSource"))
```

Set `na.rm = FALSE` to keep the missing values as their own stratum; `na_label`
names it:

```{r na-keep, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("RiskOfBias", "FundingSource"),
               na.rm = FALSE, na_label = "Not reported")
```

## `labels`: what to print on each stratum

`labels` controls the annotation inside each stratum box. The default `"none"`
leaves them bare:

```{r labels-none, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), labels = "none")
```

`"count"` adds the number of studies in each stratum:

```{r labels-count, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), labels = "count")
```

`"prop"` shows the proportion instead:

```{r labels-prop, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), labels = "prop")
```

`"both"` prints count and proportion together:

```{r labels-both, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), labels = "both")
```

## `flow_labels`: counts on the ribbons

Set `flow_labels = TRUE` to print the number of studies on each ribbon between
strata, making individual co-occurrences legible:

```{r flow-labels, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), flow_labels = TRUE)
```

## `flow_alpha`: ribbon transparency

`flow_alpha` sets the opacity of the ribbons, from 0 (invisible) to 1 (opaque).
Low values keep dense diagrams readable by letting overlaps show through:

```{r flow-alpha-low, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), flow_alpha = 0.15)
```

A high value makes the flows bold and solid:

```{r flow-alpha-high, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), flow_alpha = 0.9)
```

## `stratum_width`: width of the stratum boxes

`stratum_width` runs from 0 to 1 (fraction of the axis band). Narrow boxes give
the ribbons more room:

```{r stratum-narrow, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), stratum_width = 0.2)
```

Wide boxes emphasise the strata over the flows:

```{r stratum-wide, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome"), stratum_width = 0.8)
```

## `axis_labels`: custom axis titles

By default each axis is titled with its column name. Pass `axis_labels`, a
character vector matching `cols`, to rename them:

```{r axis-labels, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome", "AgeGroup"),
               axis_labels = c("Study design", "Outcome", "Age group"))
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose, fig.height = 5, eval = has_ggalluvial}
reviewAlluvial(studies, c("Design", "Outcome")) +
  labs(title = "Designs and outcomes", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
