---
title: "Bar charts"
description: >
  Every configuration option of reviewBar(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
has_ggfittext <- requireNamespace("ggfittext", quietly = TRUE)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewBar()` summarizes a single column as a horizontal bar chart with
frequency and percentage labels. It is the workhorse plot of the package. This
article walks through **every argument** of

```r
reviewBar(data, col, fill = "#7BB0D1", width = 0.6, sep = "\r\n",
          studlabs = FALSE, study_id = StudyID, label_space = 1.6,
          base_size = 12, na.rm = TRUE, na_label = "Not reported",
          na_in_percent = TRUE, na_last = FALSE)
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Bars are
ordered by frequency, and each is annotated with its count and percentage.

```{r default}
reviewBar(studies, Design)
```

## `col`: the column to summarize

Any categorical column works. Multi-value columns (cells holding several
values) are split first — see [`sep`](#sep-multi-value-separator) below.

```{r col-setting}
reviewBar(studies, Setting)
```

## `fill`: bar color

A single hex color paints every bar:

```{r fill-single}
reviewBar(studies, Design, fill = "#59a14f")
```

Use one of the package [`PALETTE`](../reference/PALETTE.html) colors:

```{r fill-palette}
reviewBar(studies, Design, fill = PALETTE[4])
```

Passing a **vector** of colors maps one color per bar (recycled if shorter than
the number of categories):

```{r fill-vector}
reviewBar(studies, Design, fill = PALETTE)
```

A **named** vector pins specific colors to specific categories:

```{r fill-named}
reviewBar(studies, Design,
          fill = c("RCT" = "#f16769", "Cohort" = "#7ea9c7"))
```

## `width`: bar thickness

`width` runs from 0 to 1 (fraction of the available band). Thin bars:

```{r width-thin}
reviewBar(studies, Design, width = 0.3)
```

Full-width bars:

```{r width-full}
reviewBar(studies, Design, width = 1)
```

## `sep`: multi-value separator

Some cells hold several values. In `studies`, `Outcome` uses newline
separators (`"\r\n"`, the default), so each value is counted independently:

```{r sep-default, fig.height = 3}
reviewBar(studies, Outcome)
```

If your data uses a different delimiter, set `sep`. Here we rebuild a
semicolon-separated column to demonstrate:

```{r sep-custom, fig.height = 3}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewBar(studies_semi, Outcome, sep = "; ")
```

## `studlabs` and `study_id`: label bars with study IDs

Set `studlabs = TRUE` to overlay the contributing study identifiers on each
bar. This requires the **ggfittext** package.

```{r studlabs, fig.height = 5, eval = has_ggfittext}
reviewBar(studies, Design, fill = PALETTE[2], studlabs = TRUE)
```

`study_id` selects which column supplies those identifiers (default
`StudyID`). Here we label with the author instead:

```{r studlabs-id, fig.height = 5, eval = has_ggfittext}
reviewBar(studies, Design, studlabs = TRUE, study_id = Author)
```

## `label_space`: room for the count labels

The count/percentage labels sit to the right of each bar. If a long label is
clipped, increase `label_space` (a multiplier on the x-axis headroom, default
`1.6`). Country names are long, so give them more room:

```{r label-space, fig.height = 9}
reviewBar(studies, Country, fill = PALETTE[6], label_space = 2)
```

A tighter value packs the plot horizontally:

```{r label-space-tight}
reviewBar(studies, Design, label_space = 1.2)
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small}
reviewBar(studies, Design, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewBar(studies, Design, base_size = 18)
```

## Missing data: `na.rm`, `na_label`, `na_in_percent`, `na_last`

These four arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

By default `na.rm = TRUE` drops missing rows, but the percentage denominator
is still the full sample, so the shown percentages need not sum to 100%:

```{r na-default, fig.height = 3}
reviewBar(studies, FundingSource)
```

Keep the missing rows as their own category with `na.rm = FALSE`; `na_label`
sets its name:

```{r na-keep, fig.height = 3}
reviewBar(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

With `na_in_percent = FALSE` the denominator excludes missing rows, so the
reported categories sum to 100%:

```{r na-in-percent, fig.height = 3}
reviewBar(studies, FundingSource, na.rm = FALSE, na_in_percent = FALSE)
```

`na_last = TRUE` forces the missing-value bar to the end (bottom) regardless of
its frequency:

```{r na-last, fig.height = 3}
reviewBar(studies, FundingSource, na.rm = FALSE, na_last = TRUE)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewBar(studies, Design, fill = "#59a14f") +
  labs(title = "Study designs", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
