---
title: "Histograms"
description: >
  Every configuration option of reviewHistogram(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewHistogram()` bins a single **numeric** column and plots the frequency of
each bin. It optionally stacks the bars by a categorical grouping column. It is
the counterpart of [`reviewBar()`](bar.html) for continuous variables. This
article walks through **every argument** of

```r
reviewHistogram(data, col, fill_by = NULL, bins = 30, binwidth = NULL,
                fill = "#7BB0D1", colors = PALETTE, sep = "\r\n",
                base_size = 12, na.rm = TRUE, na_label = "Not reported")
```

## Default

Pass the data frame and a numeric column. Column names may be bare or quoted.
Values are split into `bins` (30 by default) and the height of each bar is the
number of studies falling in that bin.

```{r default}
reviewHistogram(studies, SampleSize)
```

## `col`: the column to bin

Any numeric column works. `Year` is a natural candidate — it shows the temporal
spread of the reviewed studies.

```{r col-year}
reviewHistogram(studies, Year)
```

Non-numeric columns are rejected: `reviewHistogram()` is for continuous
variables. Use [`reviewBar()`](bar.html) for categorical ones.

## `fill_by`: stack by a grouping column

Supply a categorical column to `fill_by` and the histogram becomes a **stacked**
histogram: each bin is split into colored segments, one per group. This shows
how the composition of a numeric variable varies across categories.

```{r fill-by-design}
reviewHistogram(studies, SampleSize, fill_by = Design)
```

`InterventionType` works just as well:

```{r fill-by-intervention}
reviewHistogram(studies, SampleSize, fill_by = InterventionType)
```

When `fill_by` is set the bars are colored from the [`colors`](#colors-the-palette-for-grouped-bars)
palette and the single [`fill`](#fill-bar-color) color is ignored.

## `bins`: number of bins

`bins` controls how finely the range is divided (default `30`). A small value
gives broad, coarse bars:

```{r bins-small}
reviewHistogram(studies, SampleSize, bins = 6)
```

A large value gives many narrow bars, revealing finer structure:

```{r bins-large}
reviewHistogram(studies, SampleSize, bins = 40)
```

## `binwidth`: fixed bin width

Set `binwidth` to fix the width of each bin in the units of the data. This
**overrides** `bins` when supplied. Here each bar spans 100 participants:

```{r binwidth}
reviewHistogram(studies, SampleSize, binwidth = 100)
```

Because `binwidth` takes precedence, `bins` has no effect once it is set:

```{r binwidth-overrides}
reviewHistogram(studies, SampleSize, binwidth = 250, bins = 40)
```

## `fill`: bar color

When `fill_by` is `NULL`, a single hex color paints every bar:

```{r fill-single}
reviewHistogram(studies, SampleSize, fill = "#59a14f")
```

Use one of the package [`PALETTE`](../reference/PALETTE.html) colors:

```{r fill-palette}
reviewHistogram(studies, SampleSize, fill = PALETTE[4])
```

## `colors`: the palette for grouped bars

`colors` supplies the fill colors cycled across groups when `fill_by` is set
(it has no effect otherwise). The default is the package `PALETTE`:

```{r colors-palette}
reviewHistogram(studies, SampleSize, fill_by = Design, colors = PALETTE)
```

Pass your own vector to recolor the groups; it is recycled if shorter than the
number of categories:

```{r colors-custom}
reviewHistogram(studies, SampleSize, fill_by = Design,
                colors = c("#f16769", "#7ea9c7", "#edc948"))
```

A **named** vector pins specific colors to specific groups:

```{r colors-named}
reviewHistogram(studies, SampleSize, fill_by = Design,
                colors = c("RCT" = "#f16769", "Cohort" = "#7ea9c7"))
```

## `sep`: multi-value separator

`sep` only matters when `fill_by` holds multi-value cells (several values per
row, newline-separated by default). Numeric columns are single-valued, and the
usual grouping columns (`Design`, `InterventionType`) are too, so this argument
is rarely needed here. It mirrors the [`sep` of `reviewBar()`](bar.html#sep-multi-value-separator):
change it if your grouping column uses a different delimiter, e.g.
`sep = "; "`.

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small}
reviewHistogram(studies, SampleSize, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewHistogram(studies, SampleSize, base_size = 18)
```

## Missing data: `na.rm` and `na_label`

These two arguments control how `NA` cells are treated. We use a column that
actually has missing values — `FollowUpWeeks`.

By default `na.rm = TRUE` silently drops rows whose value in `col` (and, when
grouping, in `fill_by`) is missing:

```{r na-default}
reviewHistogram(studies, FollowUpWeeks)
```

`na_label` names the missing-value group when a `fill_by` column contains `NA`
and `na.rm = FALSE`. It labels those rows as their own stacked segment rather
than dropping them:

```{r na-label}
reviewHistogram(studies, SampleSize, fill_by = FundingSource,
                na.rm = FALSE, na_label = "Not reported")
```

Note that `na_label` applies to missing values in `fill_by`, not in `col`:
numeric bins have no bar for `NA`.

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewHistogram(studies, SampleSize, fill = "#59a14f") +
  labs(title = "Sample sizes", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
