---
title: "World maps"
description: >
  Every configuration option of reviewMap(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 10, fig.height = 5, out.width = "100%",
  message = FALSE, warning = FALSE
)
has_maps <- requireNamespace("maps", quietly = TRUE)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewMap()` draws a world choropleth, shading each country by the number of
studies conducted there. It is the geographic counterpart to `reviewBar()`.
This article walks through **every argument** of

```r
reviewMap(data, country_col = Country, sep = "\r\n", fill = "#7BB0D1",
          base_size = 12, na.rm = TRUE)
```

The function relies on the **maps** package to supply country polygons; all
map examples below are skipped when it is not installed.

## Default

Pass the data frame — `country_col` defaults to `Country`. Countries are
counted and the world is shaded from a light tint (few studies) to `fill` (many
studies). Unvisited countries stay a neutral grey.

```{r default, eval = has_maps}
reviewMap(studies)
```

## `country_col`: the country column

The default is `Country`, a multi-value column in `studies` where each cell
lists every country a study covers, separated by newlines. Column names may be
bare or quoted; passing it explicitly is equivalent to the default:

```{r country-col, eval = has_maps}
reviewMap(studies, country_col = Country)
```

## `sep`: multi-value separator

A single cell may name several countries. In `studies`, `Country` uses newline
separators (`"\r\n"`, the default), so a multi-country study contributes one
count to each of its countries:

```{r sep-default, eval = has_maps}
reviewMap(studies, sep = "\r\n")
```

If your data uses a different delimiter, set `sep`. Here we rebuild a
semicolon-separated column to demonstrate:

```{r sep-custom, eval = has_maps}
studies_semi <- studies
studies_semi$Country <- gsub("\r\n", "; ", studies_semi$Country)
reviewMap(studies_semi, sep = "; ")
```

## `fill`: high end of the color gradient

`fill` sets the **high** color of the gradient (the low end is a fixed light
tint of the same hue). A single hex color works:

```{r fill-hex, eval = has_maps}
reviewMap(studies, fill = "#59a14f")
```

Use one of the package [`PALETTE`](../reference/PALETTE.html) colors:

```{r fill-palette7, eval = has_maps}
reviewMap(studies, fill = PALETTE[7])
```

A different palette entry shifts the whole ramp:

```{r fill-palette6, eval = has_maps}
reviewMap(studies, fill = PALETTE[6])
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small, eval = has_maps}
reviewMap(studies, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large, eval = has_maps}
reviewMap(studies, base_size = 18)
```

## `na.rm`: handling missing countries

`reviewMap()` supports only `na.rm` (there is no `na_label`, `na_in_percent`,
or `na_last`). By default `na.rm = TRUE` drops rows with a missing or empty
country before counting. The `studies` dataset has no missing countries, so we
build a small frame that does:

```{r na-data}
studies_na <- data.frame(
  StudyID = c("S1", "S2", "S3", "S4"),
  Country = c("Spain", "Germany", NA, ""),
  stringsAsFactors = FALSE
)
```

With the default, the two studies without a country simply do not contribute:

```{r na-default, eval = has_maps}
reviewMap(studies_na)
```

With `na.rm = FALSE`, the missing rows are retained and grouped under an
`"Unknown"` region. `"Unknown"` matches no country polygon, so it never appears
on the map, but it *is* counted — this keeps the totals consistent with the raw
data even though it changes nothing visible here:

```{r na-keep, eval = has_maps}
reviewMap(studies_na, na.rm = FALSE)
```

## Automatic country-alias resolution

Country names in your data need not match the exact spelling used by
`maps::map_data("world")`. Common aliases are resolved automatically — for
example `"United States"` -> `USA`, `"United Kingdom"` -> `UK`, and
`"Czechia"` -> `Czech Republic`. All three shade correctly below:

```{r alias-data}
studies_alias <- data.frame(
  StudyID = c("S1", "S2", "S3", "S4", "S5"),
  Country = c("United States", "United States",
              "United Kingdom", "Czechia", "Spain"),
  stringsAsFactors = FALSE
)
```

```{r alias-map, eval = has_maps}
reviewMap(studies_alias, fill = PALETTE[3])
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose, eval = has_maps}
reviewMap(studies, fill = "#59a14f") +
  labs(title = "Study locations", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
