---
title: "Coding matrices"
description: >
  Every configuration option of reviewMatrix(), the study-by-criteria evidence
  matrix.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 9, fig.height = 6, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewMatrix()` draws an **evidence / coding matrix**: one row per study, one
column per criterion, and a tile wherever a study addresses a criterion. The
tile **fill** encodes an optional per-study attribute (such as publication
type) and the **letter** inside each tile is that cell's own value — a coding
level such as `F`/`P`/`M`. Column headers can carry the number of studies
addressing each criterion. This article walks through **every argument** of

```r
reviewMatrix(data, cols, color_by = NULL, study_id = StudyID,
             levels = NULL, colors = PALETTE, show_counts = TRUE,
             base_size = 12, label_wrap = 20,
             empty_fill = "#FCFCE6", tile_color = "white")
```

## Input shape

The input is **wide**: one row per study, a study-id column, an optional
grouping column for the fill, and one column per criterion holding the cell
code (or `NA`/`""` where the study does not address that criterion). The bundled
`studies` data carries a block of methodological **reporting criteria** coded
exactly this way — `"F"` full, `"P"` partial, `"M"` only mentioned, `NA` not
addressed — plus a `PubType` column to colour by:

```{r data}
criteria <- c("Randomization", "Blinding", "SampleJustification",
              "AttritionReported", "EthicsApproval", "Preregistration",
              "EffectSize", "LimitationsDiscussed")
studies[1:6, c("StudyID", "PubType", "Randomization", "EthicsApproval")]
```

To keep the figures legible we plot a subset of studies; pass the full data
frame to show them all.

```{r subset}
sel <- studies[1:22, ]
```

## Default

Pass the data frame and the vector of criterion columns. Each coded cell becomes
a tile carrying its letter; blank cells stay empty. Here we also colour by
`PubType` and describe the level codes — the two most common additions:

```{r default, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             levels = c(F = "Full", P = "Partial", M = "Mention"))
```

Randomization and blinding are trial-only items, so they are blank for the
non-RCT studies — a faithful, if sparse, corner of the matrix.

## `cols`: the criterion columns

`cols` is a character vector naming the columns to place on the x-axis, in the
order given. Pass a subset to focus on particular criteria:

```{r cols-subset, fig.height = 6}
reviewMatrix(sel, c("EthicsApproval", "Preregistration", "EffectSize"),
             color_by = "PubType")
```

## `color_by`: tile fill

`color_by` names a per-study column mapped to the tile fill, and the studies are
grouped by it so each category clusters together. Omit it (the default `NULL`)
to fill every tile with a single colour:

```{r color-null, fig.height = 6.5}
reviewMatrix(sel, criteria)
```

## `study_id`: row labels

`study_id` selects the column shown on the y-axis (default `StudyID`). Label
with the author instead:

```{r study-id, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType", study_id = Author)
```

## `levels`: the Level legend

`levels` is a named vector mapping each cell code to a human-readable
description. Its names also fix the legend order. Supply it to add a labelled
**Level** legend:

```{r levels, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             levels = c(F = "Full", P = "Partial", M = "Mention"))
```

Leave it `NULL` and the legend falls back to the bare codes found in the data.

## `colors`: fill palette

`colors` supplies the fill colours for the `color_by` categories (default
[`PALETTE`](../reference/PALETTE.html)). A custom vector is matched to the
categories in order:

```{r colors-vec, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             colors = c("#e15759", "#4e79a7", "#59a14f", "#f28e2b"))
```

A **named** vector pins specific colours to specific categories:

```{r colors-named, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             colors = c(Journal = "#4e79a7", Conference = "#e15759",
                        Preprint = "#b07aa1", Report = "#f28e2b"))
```

## `show_counts`: counts in headers

By default each column header gains `" (N=k)"`, where `k` is the number of
studies addressing that criterion. Turn it off for bare names:

```{r show-counts, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType", show_counts = FALSE)
```

## `base_size`: overall scaling

A single knob scales text and elements proportionally:

```{r base-size, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType", base_size = 16)
```

## `label_wrap`: wrap long axis labels

Long criterion or study labels wrap onto multiple lines once they exceed
`label_wrap` characters (default `20`). Lower it to wrap sooner:

```{r label-wrap, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType", label_wrap = 8)
```

## `empty_fill` and `tile_color`

`empty_fill` is the colour of the background grid behind empty cells, and
`tile_color` is the border drawn between every tile. Together they control how
strongly the matrix grid reads:

```{r grid, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             empty_fill = "grey95", tile_color = "grey80")
```

## Composing with ggplot2

`reviewMatrix()` returns a plain ggplot, so you can keep adding layers and
labels with `+`:

```{r compose, fig.height = 6.5}
reviewMatrix(sel, criteria, color_by = "PubType",
             levels = c(F = "Full", P = "Partial", M = "Mention")) +
  labs(title = "Reporting-criteria matrix",
       subtitle = "How fully each study reports each item",
       x = "Reporting criteria", y = "Study")
```
