---
title: "Co-occurrence heatmaps"
description: >
  Every configuration option of reviewOverlap(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewOverlap()` cross-tabulates two columns and draws the result as a tile
heatmap: each tile counts how many studies share that particular combination of
`col1` (x-axis) and `col2` (y-axis). Tile color runs from light grey (few
studies) to `fill` (many). This article walks through **every argument** of

```r
reviewOverlap(data, col1, col2, sep = "\r\n", fill = "#7BB0D1",
              base_size = 12, na.rm = TRUE, na_label = "Not reported",
              studlabs = FALSE, study_id = StudyID)
```

## Default

Pass the data frame and two columns. Column names may be bare or quoted. Every
combination that occurs at least once gets a tile, annotated with its count.

```{r default}
reviewOverlap(studies, Design, Setting)
```

## `col1`: the x-axis column

`col1` is mapped to the horizontal axis. Prefer columns with a handful of
categories for a readable grid. Here `RiskOfBias` (three levels) sits on x:

```{r col1}
reviewOverlap(studies, RiskOfBias, Setting)
```

## `col2`: the y-axis column

`col2` is mapped to the vertical axis. Either column may be **multi-value** — a
cell holding several values split by `sep`. In `studies`, `AgeGroup` lists one
or more age bands per study, so each band is counted independently:

```{r col2-multi, fig.height = 4}
reviewOverlap(studies, Design, AgeGroup)
```

When both columns have several categories the grid grows; the pairing below
crosses intervention type with age group:

```{r col2-multi2, fig.height = 4}
reviewOverlap(studies, InterventionType, AgeGroup)
```

## `sep`: multi-value separator

Multi-value cells are split on `sep` (default `"\r\n"`) before counting.
`AgeGroup` uses the default, so no extra argument is needed above. If your data
uses a different delimiter, set `sep`. Here we rebuild a semicolon-separated
column to demonstrate:

```{r sep-custom, fig.height = 4}
studies_semi <- studies
studies_semi$AgeGroup <- gsub("\r?\n", "; ", studies_semi$AgeGroup)
reviewOverlap(studies_semi, Design, AgeGroup, sep = "; ")
```

## `fill`: high-end gradient color

The count gradient runs from a fixed light grey (low counts) up to `fill`
(high counts). A single hex color sets that high end:

```{r fill-hex}
reviewOverlap(studies, Design, Setting, fill = "#59a14f")
```

Use one of the package [`PALETTE`](../reference/PALETTE.html) colors — a green
high end:

```{r fill-palette-3}
reviewOverlap(studies, Design, Setting, fill = PALETTE[3])
```

Or a warm orange high end:

```{r fill-palette-7}
reviewOverlap(studies, Design, Setting, fill = PALETTE[7])
```

## `base_size`: overall text and element scaling

A single knob scales all text, tile borders, and spacing proportionally.
Smaller, for multi-panel figures:

```{r base-size-small}
reviewOverlap(studies, Design, Setting, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewOverlap(studies, Design, Setting, base_size = 18)
```

## Missing data: `na.rm` and `na_label`

These two arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `RiskOfBias` has a few unreported
studies.

By default `na.rm = TRUE` drops rows missing either column, so no tile
represents them:

```{r na-default}
reviewOverlap(studies, RiskOfBias, Setting)
```

Keep the missing values as their own row/column with `na.rm = FALSE`;
`na_label` names the resulting category:

```{r na-keep}
reviewOverlap(studies, RiskOfBias, Setting, na.rm = FALSE)
```

`na_label` sets a custom label for those cells:

```{r na-label}
reviewOverlap(studies, RiskOfBias, Setting,
              na.rm = FALSE, na_label = "Unclear")
```

## `studlabs` and `study_id`: label tiles with study IDs

Set `studlabs = TRUE` to write the contributing study identifiers inside each
tile, beneath the count:

```{r studlabs, fig.height = 5}
reviewOverlap(studies, Design, RiskOfBias, studlabs = TRUE)
```

`study_id` selects which column supplies those identifiers (default
`StudyID`). Here we label with the author instead:

```{r studlabs-id, fig.height = 5}
reviewOverlap(studies, Design, RiskOfBias,
              studlabs = TRUE, study_id = Author)
```

## `label_wrap`: wrap long axis labels

Category names on the axes are wrapped onto multiple lines once they exceed
`label_wrap` characters (default `15`), so long labels stay readable instead of
running off the plot. Lower the width to wrap more aggressively:

```{r label-wrap, fig.height = 5}
reviewOverlap(studies, AnalysisApproach, Setting, label_wrap = 10)
```

Set `label_wrap = NULL` (or `Inf`) to disable wrapping and keep every label on
a single line:

```{r label-wrap-off, fig.height = 5}
reviewOverlap(studies, AnalysisApproach, Setting, label_wrap = NULL)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewOverlap(studies, Design, Setting, fill = "#59a14f") +
  labs(title = "Design by setting", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
