---
title: "Pie and donut charts"
description: >
  Every configuration option of reviewPie(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewPie()` summarizes a single column as a donut (default) or pie chart, with
each slice annotated by its count and percentage. It is the categorical
counterpart to [`reviewBar()`](bar.html). This article walks through **every
argument** of

```r
reviewPie(data, col, sep = "\r\n", colors = PALETTE, donut = TRUE,
          study_id = StudyID, base_size = 12, na.rm = TRUE,
          na_label = "Not reported", na_in_percent = TRUE, na_last = FALSE)
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Slices are
sized by frequency and labelled with their count and percentage. By default you
get a donut (a pie with a hollow center).

```{r default}
reviewPie(studies, Design)
```

## `col`: the column to summarize

Any categorical column works. Multi-value columns (cells holding several
values) are split first — see [`sep`](#sep-multi-value-separator) below.

```{r col-outcome}
reviewPie(studies, Outcome)
```

## `sep`: multi-value separator

Some cells hold several values. In `studies`, `Outcome` uses newline separators
(`"\r\n"`, the default), so each value is counted independently. If your data
uses a different delimiter, set `sep`. Here we rebuild a semicolon-separated
column to demonstrate:

```{r sep-custom}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewPie(studies_semi, Outcome, sep = "; ")
```

## `colors`: slice fill colors

By default slices are filled from the package [`PALETTE`](../reference/PALETTE.html),
cycled across categories:

```{r colors-palette}
reviewPie(studies, Design, colors = PALETTE)
```

Passing a **custom vector** of colors maps one color per slice (recycled if
shorter than the number of categories):

```{r colors-vector}
reviewPie(studies, Design, colors = c("#59a14f", "#f28e2b", "#4e79a7"))
```

A **named** vector pins specific colors to specific categories, regardless of
slice order:

```{r colors-named}
reviewPie(studies, Design,
          colors = c("RCT" = "#f16769", "Cohort" = "#7ea9c7",
                     "Case-control" = "#59a14f"))
```

## `donut`: donut vs. full pie

`donut = TRUE` (the default) leaves a hollow center. Set `donut = FALSE` for a
solid, filled pie:

```{r donut-false}
reviewPie(studies, Design, donut = FALSE)
```

The donut form again for contrast:

```{r donut-true}
reviewPie(studies, Design, donut = TRUE)
```

## `study_id`: the identifier column

`study_id` names the column of study identifiers used when counting rows
(default `StudyID`). It rarely needs changing, but if your identifiers live in a
differently named column, point `reviewPie()` at it. Here `Author` serves as the
identifier:

```{r study-id}
reviewPie(studies, Design, study_id = Author)
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small}
reviewPie(studies, Design, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewPie(studies, Design, base_size = 18)
```

## Missing data: `na.rm`, `na_label`, `na_in_percent`, `na_last`

These four arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

By default `na.rm = TRUE` drops missing rows, but the percentage denominator is
still the full sample, so the shown percentages need not sum to 100%:

```{r na-default}
reviewPie(studies, FundingSource)
```

Keep the missing rows as their own slice with `na.rm = FALSE`; `na_label` sets
its name:

```{r na-keep}
reviewPie(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

With `na_in_percent = FALSE` the denominator excludes missing rows, so the
reported categories sum to 100%:

```{r na-in-percent}
reviewPie(studies, FundingSource, na.rm = FALSE, na_in_percent = FALSE)
```

`na_last = TRUE` forces the missing-value slice to the end regardless of its
frequency:

```{r na-last}
reviewPie(studies, FundingSource, na.rm = FALSE, na_last = TRUE)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewPie(studies, Design, colors = PALETTE) +
  labs(title = "Study designs", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
