---
title: "Stacked bar charts"
description: >
  Every configuration option of reviewStackedBar(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewStackedBar()` cross-tabulates two columns: `col` defines the primary
category (one bar each) and `group` is split within every bar as the fill.
It answers questions like "how does risk of bias break down by study design?".
This article walks through **every argument** of

```r
reviewStackedBar(data, col, group, position = c("fill", "stack"),
                 fill = PALETTE, width = 0.7, sep = "\r\n",
                 study_id = StudyID, base_size = 12, na.rm = TRUE,
                 na_label = "Not reported", labels = TRUE)
```

## Default

Pass the data frame, the primary column, and the grouping column. Column names
may be bare or quoted. Each `col` category becomes one bar, split by `group`;
by default (`position = "fill"`) every bar is scaled to 100% so you read
proportions, and each segment is annotated with its count and percentage.

```{r default}
reviewStackedBar(studies, Design, RiskOfBias)
```

## `col`: the primary category

`col` sets one bar per level. Any categorical column works; multi-value cells
are split first — see [`sep`](#sep-multi-value-separator) below. Here we make
one bar per intervention type:

```{r col-intervention}
reviewStackedBar(studies, InterventionType, OpenAccess)
```

## `group`: the splitting (fill) variable

`group` is stacked within each bar and drives the legend. Swapping `col` and
`group` re-frames the same cross-tabulation — now one bar per risk-of-bias
level, split by design:

```{r group-swap, fig.height = 3}
reviewStackedBar(studies, RiskOfBias, Design)
```

## `position`: proportions vs counts

`position = "fill"` (the default) scales each bar to 100%, so bars are
comparable regardless of how many studies fall in each `col` category:

```{r position-fill}
reviewStackedBar(studies, Design, RiskOfBias, position = "fill")
```

`position = "stack"` keeps the raw counts, so bar height reflects the number of
studies and the labels report frequencies:

```{r position-stack}
reviewStackedBar(studies, Design, RiskOfBias, position = "stack")
```

## `fill`: segment colors

The default is the package [`PALETTE`](../reference/PALETTE.html), applied
across the `group` levels:

```{r fill-palette}
reviewStackedBar(studies, Design, RiskOfBias, fill = PALETTE)
```

Pass a **vector** of colors to map one color per `group` level (recycled if
shorter than the number of levels):

```{r fill-vector}
reviewStackedBar(studies, Design, RiskOfBias,
                 fill = c("#59a14f", "#edc948", "#e15759"))
```

A **named** vector pins specific colors to specific `group` levels, independent
of their order — useful to keep, say, "High" red across figures:

```{r fill-named}
reviewStackedBar(studies, Design, RiskOfBias,
                 fill = c("Low" = "#59a14f",
                          "Moderate" = "#edc948",
                          "High" = "#e15759"))
```

## `width`: bar thickness

`width` runs from 0 to 1 (fraction of the available band). Thin bars:

```{r width-thin}
reviewStackedBar(studies, Design, RiskOfBias, width = 0.3)
```

Full-width bars:

```{r width-full}
reviewStackedBar(studies, Design, RiskOfBias, width = 1)
```

## `sep`: multi-value separator

Either `col` or `group` may hold several values per cell. In `studies`,
`Outcome` uses newline separators (`"\r\n"`, the default), so a study
reporting multiple outcomes contributes to each. Here `Outcome` is the primary
category, split by design:

```{r sep-default, fig.height = 3}
reviewStackedBar(studies, Outcome, Design)
```

If your data uses a different delimiter, set `sep`. Here we rebuild a
semicolon-separated column to demonstrate:

```{r sep-custom, fig.height = 3}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewStackedBar(studies_semi, Outcome, Design, sep = "; ")
```

## `study_id`: which column identifies a study

`study_id` names the column used to identify studies when de-duplicating and
counting (default `StudyID`). It matters for multi-value columns, where the same
study can appear in several segments. Pointing it at `Author` uses that column
as the identifier instead:

```{r study-id, fig.height = 3}
reviewStackedBar(studies, Outcome, Design, study_id = Author)
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small}
reviewStackedBar(studies, Design, RiskOfBias, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewStackedBar(studies, Design, RiskOfBias, base_size = 18)
```

## `na.rm` and `na_label`: missing data in the group

This function exposes two missing-data controls, both acting on the `group`
column. By default `na.rm = TRUE` drops rows with a missing `group` value.
`OpenAccess` has some unreported entries, which vanish here:

```{r na-default}
reviewStackedBar(studies, InterventionType, OpenAccess)
```

Set `na.rm = FALSE` to keep those rows as their own segment; `na_label` sets its
name:

```{r na-keep}
reviewStackedBar(studies, InterventionType, OpenAccess,
                 na.rm = FALSE, na_label = "Not reported")
```

The same pair works for any group with gaps, such as `RiskOfBias`:

```{r na-riskofbias}
reviewStackedBar(studies, Design, RiskOfBias,
                 na.rm = FALSE, na_label = "Unclear")
```

## `labels`: in-segment annotations

`labels = TRUE` (the default) prints the count/percentage inside each segment.
Turn it off for a cleaner look, leaving the legend to carry the meaning:

```{r labels-off}
reviewStackedBar(studies, Design, RiskOfBias, labels = FALSE)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewStackedBar(studies, Design, RiskOfBias) +
  labs(title = "Risk of bias by study design",
       subtitle = "n = 50 studies", fill = "Risk of bias",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
