---
title: "Summary tables"
description: >
  Every configuration option of reviewTable(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
data(studies)
```

`reviewTable()` summarizes a single column as a formatted table listing each
category, its contributing studies, frequency, and percentage. It returns a
[`gt`](https://gt.rstudio.com) object that renders inline in HTML and can be
exported to LaTeX. This article walks through **every argument** of

```r
reviewTable(data, col, sep = "\r\n", study_id = StudyID, latex = FALSE,
            cite = FALSE, na.rm = TRUE, na_label = "Not reported",
            na_in_percent = TRUE, na_last = FALSE)
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Each row
is a distinct category, with its contributing study IDs, count, and percentage.

```{r default}
reviewTable(studies, Design)
```

## `col`: the column to summarize

Any categorical column works. Multi-value columns (cells holding several
values) are split first — see [`sep`](#sep-multi-value-separator) below.
`Outcome` stores several outcomes per study, so each value is counted
independently and its study list is the union of studies reporting it:

```{r col-outcome}
reviewTable(studies, Outcome)
```

## `sep`: multi-value separator

Multi-value cells are split on `sep` before counting. In `studies`, `Outcome`
uses newline separators (`"\r\n"`, the default). If your data uses a different
delimiter, set `sep`. Here we rebuild a semicolon-separated column to
demonstrate:

```{r sep-custom}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewTable(studies_semi, Outcome, sep = "; ")
```

## `study_id`: which column lists the contributing studies

By default the studies column is filled from `StudyID`. Point `study_id` at any
other identifier column to list studies differently — here, by author:

```{r study-id-author}
reviewTable(studies, Design, study_id = Author)
```

## `latex`: emit LaTeX instead of a gt table

With `latex = TRUE`, `reviewTable()` returns a LaTeX representation of the table
(via [`gt::as_latex()`](https://gt.rstudio.com/reference/as_latex.html)) rather
than an HTML `gt` object. The result is a character string of class
`litreview_latex`; in a knitted document it is emitted verbatim. Below we show
the generated LaTeX source as text so it is visible in this HTML article:

```{r latex, results = "asis"}
tex <- reviewTable(studies, Design, latex = TRUE)
cat("```latex", tex, "```", sep = "\n")
```

Drop that block into a `.tex` document (it requires the `booktabs` package) to
typeset the table.

## `cite`: render studies as `\cite{}` keys

`cite = TRUE` treats the `study_id` values as BibTeX citation keys and wraps
each row's study list in a single `\cite{...}` command — ideal for a
manuscript. It is most useful with `study_id = bibKey` (the dataset's BibTeX
keys) and `latex = TRUE`:

```{r cite, results = "asis"}
tex <- reviewTable(studies, Design, study_id = bibKey,
                   latex = TRUE, cite = TRUE)
cat("```latex", tex, "```", sep = "\n")
```

Each `Studies` cell now holds `\cite{author6_2021,author17_2019,...}`, which
resolves against your `.bib` file at compile time.

## Embedding in a LaTeX or R Markdown document

When your document compiles to PDF — an `.Rnw` (Sweave) file, or an R Markdown
document with a LaTeX/PDF output format — call `reviewTable(..., latex = TRUE)`
in a chunk and the table typesets in place. The returned object carries a
`knit_print` method that emits the LaTeX verbatim, so no extra chunk options are
needed; in Sweave, the chunk option `results = "asis"` achieves the same.

```{r embed, eval = FALSE}
reviewTable(studies, Design, study_id = bibKey, latex = TRUE, cite = TRUE)
```

That call expands to a complete `table` environment (`\begin{table}` …
`\end{table}`) — the same source shown under [`latex`](#latex-emit-latex-instead-of-a-gt-table)
and [`cite`](#cite-render-studies-as-cite-keys) above. Two preamble
requirements:

- The rules use **booktabs** (`\toprule`, `\midrule`, `\bottomrule`), so load
  `\usepackage{booktabs}`.
- With `cite = TRUE`, load a citation package (`\usepackage{natbib}` or
  `\usepackage{biblatex}`) and make sure the keys in your `study_id` column
  exist in the bibliography, so each `\cite{...}` resolves at compile time.

If your citation keys live in a column other than `StudyID`, point `study_id`
at it — for example `study_id = bibKey` above, or any custom key column:

```{r embed-custom, results = "asis"}
refs <- data.frame(
  RefKey = c("smith2020", "doe2019", "lee2021", "smith2020"),
  Design = c("RCT", "Cohort", "RCT", "Case-control"),
  stringsAsFactors = FALSE
)
tex <- reviewTable(refs, Design, study_id = RefKey, latex = TRUE, cite = TRUE)
cat("```latex", tex, "```", sep = "\n")
```

## Missing data: `na.rm`, `na_label`, `na_in_percent`, `na_last`

These four arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

### `na.rm`

By default `na.rm = TRUE` drops missing rows, but the percentage denominator is
still the full sample, so the shown percentages need not sum to 100%:

```{r na-rm-default}
reviewTable(studies, FundingSource)
```

Set `na.rm = FALSE` to keep the missing rows as their own category:

```{r na-rm-keep}
reviewTable(studies, FundingSource, na.rm = FALSE)
```

### `na_label`

When missing rows are kept, `na_label` names that category (default
`"Not reported"`):

```{r na-label}
reviewTable(studies, FundingSource, na.rm = FALSE, na_label = "Unspecified")
```

### `na_in_percent`

With `na_in_percent = FALSE` the denominator excludes missing rows, so the
reported categories sum to 100%:

```{r na-in-percent}
reviewTable(studies, FundingSource, na.rm = FALSE, na_in_percent = FALSE)
```

### `na_last`

`na_last = TRUE` forces the missing-value row to the end of the table
regardless of its frequency:

```{r na-last}
reviewTable(studies, FundingSource, na.rm = FALSE, na_last = TRUE)
```

## Notes

`reviewTable()` returns a [`gt`](https://gt.rstudio.com) object (or, with
`latex = TRUE`, a LaTeX character string). Because the default result is a plain
`gt` table, you can pipe it into any `gt` styling verb — `gt::tab_header()`,
`gt::cols_label()`, `gt::opt_stylize()`, and so on — before printing, or export
it with `gt::gtsave()` and `gt::as_latex()`.
