---
title: "Treemaps"
description: >
  Every configuration option of reviewTreemap(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
has_treemapify <- requireNamespace("treemapify", quietly = TRUE)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewTreemap()` displays category frequencies as nested rectangles whose area
is proportional to the count — a compact alternative to a bar chart when you
have many categories or a second grouping level. It requires the
**treemapify** package. This article walks through **every argument** of

```r
reviewTreemap(data, col, color_by = NULL, sep = "\r\n", colors = PALETTE,
              base_size = 12, na.rm = TRUE, na_label = "Not reported",
              study_id = StudyID, studlabs = FALSE, border_col = "white")
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Each
distinct value becomes one rectangle, sized by how often it occurs.

```{r default, eval = has_treemapify}
reviewTreemap(studies, Design)
```

## `col`: the column to summarize

Any categorical column works. Multi-value columns — cells holding several
values separated by [`sep`](#sep-multi-value-separator) — are split first, so
each value is counted independently. `Intervention` is such a column:

```{r col-intervention, eval = has_treemapify}
reviewTreemap(studies, Intervention)
```

`Outcome` is another multi-value column:

```{r col-outcome, eval = has_treemapify}
reviewTreemap(studies, Outcome)
```

## `color_by`: hierarchical grouping

Supplying `color_by` nests `col` inside a higher-order column and colors the
tiles by that grouping, with a visible subgroup border. The classic pairing is
interventions grouped by their broader type:

```{r color-by, fig.height = 5, eval = has_treemapify}
reviewTreemap(studies, Intervention, color_by = InterventionType)
```

Without `color_by` (the default `NULL`), rectangles are simply colored by
`col` itself and there is no grouping:

```{r color-by-none, fig.height = 5, eval = has_treemapify}
reviewTreemap(studies, Intervention)
```

## `sep`: multi-value separator

Multi-value cells are split on `sep` before counting. The default `"\r\n"`
matches the newline-separated cells in `studies` (used above for `Intervention`
and `Outcome`). If your data uses a different delimiter, set `sep`. Here we
rebuild a semicolon-separated column to demonstrate:

```{r sep-custom, eval = has_treemapify}
studies_semi <- studies
studies_semi$Intervention <- gsub("\r\n|\n", "; ", studies_semi$Intervention)
reviewTreemap(studies_semi, Intervention, sep = "; ")
```

## `colors`: the fill palette

By default tiles are filled from the package [`PALETTE`](../reference/PALETTE.html),
cycled to cover the categories:

```{r colors-palette, eval = has_treemapify}
reviewTreemap(studies, Design, colors = PALETTE)
```

Pass any **vector** of colors to use your own scheme (recycled if shorter than
the number of tiles):

```{r colors-vector, eval = has_treemapify}
reviewTreemap(studies, Design,
              colors = c("#59a14f", "#f28e2b", "#4e79a7", "#e15759", "#b07aa1"))
```

A **named** vector pins specific colors to specific categories:

```{r colors-named, eval = has_treemapify}
reviewTreemap(studies, Design,
              colors = c("Qualitative" = "#f16769",
                         "Cohort"      = "#7ea9c7",
                         "RCT"         = "#59a14f"))
```

When `color_by` is set, `colors` maps onto the grouping column instead:

```{r colors-group, fig.height = 5, eval = has_treemapify}
reviewTreemap(studies, Intervention, color_by = InterventionType,
              colors = PALETTE)
```

## `base_size`: overall text and element scaling

A single knob scales the text (and border thickness) proportionally. Smaller,
for multi-panel figures:

```{r base-size-small, eval = has_treemapify}
reviewTreemap(studies, Design, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large, eval = has_treemapify}
reviewTreemap(studies, Design, base_size = 18)
```

## Missing data: `na.rm` and `na_label`

These two arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

By default `na.rm = TRUE` drops missing rows entirely, so they contribute no
rectangle:

```{r na-default, eval = has_treemapify}
reviewTreemap(studies, FundingSource)
```

Keep the missing rows as their own tile with `na.rm = FALSE`; `na_label` sets
its name:

```{r na-keep, eval = has_treemapify}
reviewTreemap(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

`na_label` accepts any string:

```{r na-label, eval = has_treemapify}
reviewTreemap(studies, FundingSource, na.rm = FALSE, na_label = "Unknown funding")
```

## `study_id` and `studlabs`: label tiles with study IDs

Set `studlabs = TRUE` to print the contributing study identifiers inside each
rectangle, beneath the category name. This feature relies on the **ggfittext**
package to fit the text.

```{r studlabs, fig.height = 5, eval = requireNamespace("ggfittext", quietly = TRUE) && has_treemapify}
reviewTreemap(studies, Design, studlabs = TRUE)
```

`study_id` selects which column supplies those identifiers (default
`StudyID`). Here we label with the author instead:

```{r studlabs-id, fig.height = 5, eval = requireNamespace("ggfittext", quietly = TRUE) && has_treemapify}
reviewTreemap(studies, Design, studlabs = TRUE, study_id = Author)
```

## `border_col`: rectangle border color

Borders separate adjacent tiles. The default is `"white"`:

```{r border-white, eval = has_treemapify}
reviewTreemap(studies, Design, border_col = "white")
```

A dark border reads well on light fills:

```{r border-dark, eval = has_treemapify}
reviewTreemap(studies, Design, border_col = "grey20")
```

Pure black gives the strongest separation, and also outlines the subgroups when
`color_by` is used:

```{r border-black, fig.height = 5, eval = has_treemapify}
reviewTreemap(studies, Intervention, color_by = InterventionType,
              border_col = "black")
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose, eval = has_treemapify}
reviewTreemap(studies, Design) +
  labs(title = "Study designs", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
