---
title: "Trend charts"
description: >
  Every configuration option of reviewTrend(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewTrend()` shows how the values of one column distribute across
publication years, drawing one stacked bar per year. It answers questions like
"which study designs have grown over time?". This article walks through
**every argument** of

```r
reviewTrend(data, col, year_col = Year, sep = "\r\n", colors = PALETTE,
            base_size = 12, na.rm = TRUE, na_label = "Not reported",
            labels = c("none", "count", "percent", "both", "studies"),
            study_id = StudyID)
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Each year
becomes a stacked bar whose segments are the categories of `col`, and the
y-axis counts studies.

```{r default}
reviewTrend(studies, Design)
```

## `col`: the column to track

Any categorical column works. Here we track the analysis setting over time:

```{r col-setting}
reviewTrend(studies, Setting)
```

Multi-value columns (cells holding several values) are split first, so each
value is counted independently in its year — see
[`sep`](#sep-multi-value-separator) below. `Outcome` is such a column:

```{r col-multi}
reviewTrend(studies, Outcome)
```

## `year_col`: the year column

By default the years come from the `Year` column. If your data names it
differently, pass `year_col`. Below we rename the column and point
`reviewTrend()` at it explicitly:

```{r year-col}
studies_pubyear <- studies
studies_pubyear$PubYear <- studies_pubyear$Year
reviewTrend(studies_pubyear, Design, year_col = PubYear)
```

## `sep`: multi-value separator

Cells holding several values are split on `sep` before counting. In `studies`,
`Outcome` uses newline separators (`"\r\n"`, the default). If your data uses a
different delimiter, set `sep`. Here we rebuild a semicolon-separated column to
demonstrate:

```{r sep-custom}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewTrend(studies_semi, Outcome, sep = "; ")
```

## `colors`: the fill palette

Segments are filled by cycling through `colors`, which defaults to the package
[`PALETTE`](../reference/PALETTE.html):

```{r colors-palette}
reviewTrend(studies, Design, colors = PALETTE)
```

Pass a **custom vector** to restyle every category (recycled if shorter than
the number of categories):

```{r colors-custom}
reviewTrend(studies, Design,
            colors = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261",
                       "#e76f51", "#8ab17d"))
```

A **named** vector pins specific colors to specific categories:

```{r colors-named}
reviewTrend(studies, Design,
            colors = c("RCT" = "#f16769", "Cohort" = "#7ea9c7",
                       "Qualitative" = "#59a14f", "Cross-sectional" = "#edc948",
                       "Mixed methods" = "#b07aa1", "Case-control" = "#76b7b2"))
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing proportionally. Smaller, for
multi-panel figures:

```{r base-size-small}
reviewTrend(studies, Design, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewTrend(studies, Design, base_size = 18)
```

## Missing data: `na.rm` and `na_label`

These two arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

By default `na.rm = TRUE` drops missing rows entirely:

```{r na-default}
reviewTrend(studies, FundingSource)
```

Keep the missing rows as their own category with `na.rm = FALSE`; `na_label`
sets the name of that category:

```{r na-keep}
reviewTrend(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

## `labels`: what to print on each segment

`labels` controls the text overlaid on each stacked segment. The default is
`"none"` — clean bars with no annotation:

```{r labels-none}
reviewTrend(studies, Design, labels = "none")
```

`"count"` prints the number of studies in each segment:

```{r labels-count}
reviewTrend(studies, Design, labels = "count")
```

`"percent"` prints each segment's share **within its year**, so the segments of
one bar sum to 100%:

```{r labels-percent}
reviewTrend(studies, Design, labels = "percent")
```

`"both"` combines the two, showing count and within-year percentage:

```{r labels-both}
reviewTrend(studies, Design, labels = "both")
```

`"studies"` overlays the contributing study identifiers themselves, one per
line. This is most legible when categories are few, so give the panel some
height:

```{r labels-studies, fig.height = 6}
reviewTrend(studies, Design, labels = "studies")
```

## `study_id`: which column supplies the IDs

When `labels = "studies"`, the `study_id` column provides the identifiers
(default `StudyID`). Point it at any ID-like column — here we label with the
author instead:

```{r study-id, fig.height = 6}
reviewTrend(studies, Design, labels = "studies", study_id = Author)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewTrend(studies, Design, colors = PALETTE) +
  labs(title = "Study designs over time", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
