---
title: "UpSet plots"
description: >
  Every configuration option of reviewUpset(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 5, out.width = "100%",
  message = FALSE, warning = FALSE
)
has_ggupset <- requireNamespace("ggupset", quietly = TRUE)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewUpset()` visualizes how the values of a multi-value column **co-occur**
across studies. Each study contributes the *set* of distinct values it reports
in that column. Every observed combination of values becomes a bar whose height
counts how many studies share that exact set, and a dot matrix beneath the bars
marks which values make up each combination. This generalizes the pairwise
[`reviewOverlap()`](../reference/reviewOverlap.html) to three or more
co-occurring values. This article walks through **every argument** of

```r
reviewUpset(data, col, sep = "\r\n", study_id = StudyID, base_size = 12,
            na.rm = TRUE, na_label = "Not reported", n_intersections = 15,
            sort_by = c("freq", "degree"), fill = "#7BB0D1")
```

`reviewUpset()` requires the **ggupset** package.

## Default

Pass the data frame and a multi-value column. `Outcome` records the outcomes
each study measured, newline-separated. The tallest bars are the most common
outcome combinations; the dots beneath show which outcomes each combination
contains.

```{r default, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome)
```

`reviewUpset()` shows the two panels that answer "which combinations occur, and
how often": the intersection-size bars and the combination matrix. The optional
per-value *set-size* sidebar from some UpSet implementations is intentionally
omitted so the result stays a single, `theme_litreview`-styled ggplot you can
extend with `+`. For the individual value totals, use [`reviewBar()`](bar.html)
on the same column.

## `col`: the multi-value column to analyze

Any column whose cells may hold several values works. Here we look at how
interventions co-occur:

```{r col-intervention, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Intervention)
```

Age groups:

```{r col-agegroup, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, AgeGroup)
```

Countries — many studies span several, so their combinations are rich:

```{r col-country, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Country)
```

## `sep`: multi-value separator

Cells holding several values are split on `sep` before the sets are built. In
`studies`, the multi-value columns use newline separators (`"\r\n"`, the
default), so each value is treated independently. If your data uses a different
delimiter, set `sep`. Here we rebuild a semicolon-separated column to
demonstrate:

```{r sep-custom, fig.height = 5, eval = has_ggupset, warning = FALSE}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n|\n", "; ", studies_semi$Outcome)
reviewUpset(studies_semi, Outcome, sep = "; ")
```

## `study_id`: how studies are grouped into sets

One set is formed per study, and `study_id` names the column that identifies
studies (default `StudyID`). All rows sharing an identifier are pooled into a
single combination. Point it at a different identifier column to regroup — here
we group by `Author` instead:

```{r study-id, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, study_id = Author)
```

## `base_size`: overall text and element scaling

A single knob scales all text, bars, and matrix dots proportionally. Smaller,
for multi-panel figures:

```{r base-size-small, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, base_size = 16)
```

## Missing data: `na.rm` and `na_label`

These two arguments control how `NA` (and empty) cells are treated. By default
`na.rm = TRUE` drops missing values, so studies that report nothing for the
column contribute no set and vanish from the plot.

Set `na.rm = FALSE` to keep missing values as a distinct set member, and
`na_label` names it. A study reporting only a missing value then forms its own
single-element combination:

```{r na-keep, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

## `n_intersections`: cap on the number of combinations shown

Only the largest `n_intersections` combinations are drawn (default `15`),
keeping the plot readable when many distinct sets exist. Lower it to focus on
the few most common combinations:

```{r n-intersections, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Country, n_intersections = 6)
```

## `sort_by`: ordering the combinations

`sort_by = "freq"` (the default) orders bars by intersection size, so the most
common combinations sit on the left:

```{r sort-freq, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, sort_by = "freq")
```

`sort_by = "degree"` instead orders by the number of values in each combination,
grouping single-value sets, then pairs, then triples, and so on:

```{r sort-degree, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, sort_by = "degree")
```

## `fill`: bar and matrix-dot color

A single hex color paints every bar and every filled matrix dot:

```{r fill-single, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, fill = "#59a14f")
```

Use one of the package [`PALETTE`](../reference/PALETTE.html) colors:

```{r fill-palette, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, fill = PALETTE[4])
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose, fig.height = 5, eval = has_ggupset, warning = FALSE}
reviewUpset(studies, Outcome, fill = "#59a14f") +
  labs(title = "Outcome combinations",
       subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
