---
title: "Waffle charts"
description: >
  Every configuration option of reviewWaffle(), demonstrated on the example
  dataset.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 4, out.width = "100%",
  message = FALSE, warning = FALSE
)
```

```{r setup}
library(litReview)
library(ggplot2)
data(studies)
```

`reviewWaffle()` summarizes a single column as a grid of colored squares.
**Each square represents one study occurrence of a category**, so the block of
squares for a category is proportional to its frequency. This article walks
through **every argument** of

```r
reviewWaffle(data, col, sep = "\r\n", colors = PALETTE, ncol = 5,
             study_id = StudyID, base_size = 12, na.rm = TRUE,
             na_label = "Not reported", na_in_percent = TRUE,
             na_last = FALSE)
```

## Default

Pass the data frame and a column. Column names may be bare or quoted. Squares
are laid out row by row, ordered by frequency, and colored by category.

```{r default}
reviewWaffle(studies, Design)
```

## `col`: the column to summarize

Any categorical column works. Each square is one occurrence, so a column with
more distinct categories yields a more finely divided grid.

```{r col-setting}
reviewWaffle(studies, Setting)
```

Multi-value columns (cells holding several values) are split first, and every
value contributes its own square — see [`sep`](#sep-multi-value-separator)
below. `Outcome` is such a column, so the total number of squares exceeds the
50 studies:

```{r col-outcome, fig.height = 5}
reviewWaffle(studies, Outcome)
```

## `sep`: multi-value separator

Some cells hold several values. In `studies`, `Outcome` uses newline
separators (`"\r\n"`, the default), so each value is counted independently and
gets its own square:

```{r sep-default, fig.height = 5}
reviewWaffle(studies, Outcome)
```

If your data uses a different delimiter, set `sep`. Here we rebuild a
semicolon-separated column to demonstrate:

```{r sep-custom, fig.height = 5}
studies_semi <- studies
studies_semi$Outcome <- gsub("\r\n", "; ", studies_semi$Outcome)
reviewWaffle(studies_semi, Outcome, sep = "; ")
```

## `colors`: the fill palette

`colors` is a vector of fill colors cycled across categories. It defaults to
the package [`PALETTE`](../reference/PALETTE.html):

```{r colors-palette}
reviewWaffle(studies, Design, colors = PALETTE)
```

Pass a **custom vector** to override it. Colors are recycled if the vector is
shorter than the number of categories:

```{r colors-custom}
reviewWaffle(studies, Design,
             colors = c("#59a14f", "#f28e2b", "#4e79a7", "#e15759"))
```

A **named** vector pins specific colors to specific categories, regardless of
their order in the grid:

```{r colors-named}
reviewWaffle(studies, Design,
             colors = c("RCT" = "#f16769", "Cohort" = "#7ea9c7"))
```

## `ncol`: number of grid columns

`ncol` sets how many squares sit in each row before wrapping to the next. A
small value makes a tall, narrow grid:

```{r ncol-small, fig.height = 7}
reviewWaffle(studies, Design, ncol = 3)
```

A large value makes a short, wide grid:

```{r ncol-large, fig.height = 3}
reviewWaffle(studies, Design, ncol = 15)
```

## `study_id`: the identifier column

`study_id` names the column of study identifiers (default `StudyID`) used
internally to count occurrences. It rarely needs changing, but if your ID
lives in a different column you can point to it. The counts — and therefore the
squares — are unchanged when every row has a unique identifier:

```{r study-id}
reviewWaffle(studies, Design, study_id = Author)
```

## `base_size`: overall text and element scaling

A single knob scales all text and spacing (including the white grout between
squares) proportionally. Smaller, for multi-panel figures:

```{r base-size-small}
reviewWaffle(studies, Design, base_size = 9)
```

Larger, for slides or posters:

```{r base-size-large}
reviewWaffle(studies, Design, base_size = 18)
```

## Missing data: `na.rm`, `na_label`, `na_in_percent`, `na_last`

These four arguments control how `NA` (and empty) cells are treated. We use a
column that actually has missing values — `FundingSource`.

By default `na.rm = TRUE` drops missing rows entirely, so they contribute no
squares:

```{r na-default}
reviewWaffle(studies, FundingSource)
```

Keep the missing rows as their own category with `na.rm = FALSE`; `na_label`
sets its name and it gets its own block of squares:

```{r na-keep}
reviewWaffle(studies, FundingSource, na.rm = FALSE, na_label = "Not reported")
```

`na_in_percent` controls whether missing rows count toward the percentage
denominator used when summarizing the data. Excluding them changes the
underlying proportions (though the waffle draws one square per occurrence
either way):

```{r na-in-percent}
reviewWaffle(studies, FundingSource, na.rm = FALSE, na_in_percent = FALSE)
```

`na_last = TRUE` forces the missing-value category to sort last, so its block
of squares appears at the end of the grid regardless of its frequency:

```{r na-last}
reviewWaffle(studies, FundingSource, na.rm = FALSE, na_last = TRUE)
```

## Composing with ggplot2

Every `review*()` function returns a plain ggplot, so you can keep adding
layers, scales, and labels with `+`:

```{r compose}
reviewWaffle(studies, Design, colors = PALETTE) +
  labs(title = "Study designs", subtitle = "n = 50 studies",
       caption = "Source: example dataset") +
  theme(plot.title.position = "plot")
```
